---
title: "Reproducible Research Project 1"
author: "Hao"
date: "7/8/2021"
output: html_document
---

##Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.


This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


##Assignment Instructions

1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

##Questions

What is mean total number of steps taken per day?
What is the average daily activity pattern?
Imputing missing values
Are there differences in activity patterns between weekdays and weekends?

##Step 1
##Setting the working directory

```{r, echo = TRUE}
setwd("C:/Users/SHIH/Desktop/Reproducible Research")
if(!file.exists("./Project1")){dir.create("./Project1")}
```


##Step 2
##Code for reading and processing the data

```{r, echo = TRUE}
library(ggplot2)

activity <- read.csv("./data/activity.csv")

activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

summary(activity)
```

##Step 3
##Q1: What is mean total number of steps taken per day?

1. What is mean total number of steps taken per day?

```{r, echo = TRUE}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "darkblue", ylim = c(0,20), breaks = seq(0,25000, by=2500))

```


##Step 4
##Mean and median number of steps taken each day

Here is the mean of the total number of steps taken per day:

```{r, echo = TRUE}
mean(activity_total_steps$steps)
```

Here is the median of the total number of steps taken per day:

```{r, echo = TRUE}
median(activity_total_steps$steps)
```


##Step 5
##Q2: What is the average daily activity pattern?

2. What is the average daily activity pattern?

```{r, echo = TRUE}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="darkblue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo = TRUE}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```


##Step 6
##Q3: Imputing missing values

3. Imputing missing values

There are multiple strategies to deal with multiple value imputations.
The common strategies include:
1. Constant value imputations
2. Regression model value imputations
3. Mean/mode value substitutions
For the purpose of simplicity, in this question, I will use the mean/mode value substitution strategy to impute missing values. That is, using the mean values to substitute out the missing values in the original data set
Before doing any sort of imputation, it is helpful to understand what are the distributions of missing values by date and interval

```{r, echo = TRUE}
sum(is.na(activity$steps))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r, echo = TRUE}
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo = TRUE}
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```

## Step 7
Histogram of the total number of steps taken each day after missing values are imputed

```{r, echo = TRUE}
hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```

Here is the mean of the total number of steps taken per day:

```{r, echo = TRUE}
mean(total_steps_imputed$daily_steps)
```

Here is the median of the total number of steps taken per day:

```{r, echo = TRUE}
median(total_steps_imputed$daily_steps)
```


##Step 8
##Q4: Are there differences in activity patterns between weekdays and weekends?

4. Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, echo = TRUE}
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```


Making plot to compare the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, echo = TRUE}
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```




